Supporting Information: Thermal sensitivity across forest vertical profiles: patterns, mechanisms, and ecological implications

### 2022-06-18 response-to-reviews.Rmd

l.264-265 'their environment is less conducive to dissipating excess heat compared to upper-canopy leaves.' What evidence? This is a logical argument and should either be backed with references or labelled as a logical outcome of the conditions they have stated.

**We added an explanation to l.264-265 with references:**
**"Thus, while understory and within-canopy shade leaves can remain cooler under lower radiation, their environment is less conducive to dissipating excess heat compared to upper-canopy leaves, *because wind speeds and hence boundary layer conductance are generally smaller in the lower canopy [@robertsStomatalBoundaryLayerConductances1990; @martin_boundary_1999].***

l.269 How does lower VPD in the understory inhibit evaporative cooling, if VPD is the major driver for evaporation? I think I understand how this can be counter-intuitive, but the reasoning isn't explained so it all becomes opaque to the reader. Please explain very briefly.

**Lower VPD inhibits evaporative cooling precisely because VPD is the major driver for evaporation: Transpiration rate generally increases as VPD increases, and decreases as VPD decreases. Therefore lower VPD will mean less transpiration and hence less evaporative cooling. Perhaps the word *‘inhibits’* confused the issue; we have replaced *‘inhibits’* with *‘reduces’*. We have also modified the sentence to read, ‘In addition to lower wind speeds, higher RH (lower VPD) in the understory and inner canopy also reduces evaporative cooling (because evaporation rate generally increases with VPD), thus increasing *T~leaf~* – *T~air~* …’.” (phrase “because evaporation rate generally increases with VPD” added)**

### 2022-06-27 vertical-thermal-review.Rmd
small updates

### 1.3 Turbulent transport and wind
Vertical transport in forest canopies is dominated by turbulent transport, making vertical profiles of wind speed, water vapor, CO~2~, and temperature more difficult to predict and to generalize on the basis of theoretical first principles than radiation profiles. 
Above the canopy, turbulent transport typically mimics diffusion – that is, vertical fluxes of heat, mass and momentum are proportional to their respective vertical gradients, and to transfer coefficients [eddy diffusivities, @penmanWeatherWheatEssay1960]-- such that wind speed declines logarithmically with proximity to the canopy surface [@monteithPrinciplesEnvironmentalPhysics2013]. 
Interaction with a plant canopy also attenuates wind, causing wind speed to decline through the canopy; for example, small plant elements can rapidly dissipate momentum within dense canopies [@raupachAveragingProceduresFlow1982; @baldocchiSpectralLagcorrelationAnalysis1988].
Within canopies, however, transport often fails to mimic diffusion – e.g., transport of momentum [@shawSecondaryWindSpeed1977] or heat [@raupachLagrangianAnalysisScalar1987] may occur in the direction opposite to that predicted by gradients of wind speed or temperature, respectively – making vertical wind profiles difficult to generalize from first principles [@denmeadScalarTransportPlant1987; @RaupachM.R1989Ttip; @meyersModellingPlantCanopy1987; @katulModelingCO2Sources1999; @harmanSimpleUnifiedTheory2007].
Such counter-gradient transport may arise from the intermittent generation, at the canopy surface, of large wakes or coherent eddy structures that periodically dip down through the canopy, gathering packets of warmed, humidified and CO~2~-enriched or depleted air from beneath and within the canopy and flushing them to the atmosphere above [@finniganTurbulenceWavingWheat1979; @baldocchiTraceGasExchange1991]. 
These "sweep-eject" events can couple understory conditions more directly to conditions above the canopy than beneath.

>>>>

### 1.3 Turbulent transport and wind
Vertical transport in forest canopies is dominated by turbulent transport, making vertical profiles of wind speed, water vapor, CO~2~, and temperature more difficult to predict and to generalize on the basis of theoretical first principles than radiation profiles. 
Above the canopy, turbulent transport typically mimics diffusion – that is, vertical fluxes of heat, mass and momentum are proportional to their respective vertical gradients, and to transfer coefficients [eddy diffusivities, @penmanWeatherWheatEssay1960]-- such that wind speed declines logarithmically with proximity to the canopy surface [@monteithPrinciplesEnvironmentalPhysics2013]. 
Interaction with a plant canopy also attenuates wind, causing wind speed to decline through the canopy; for example, small plant elements can rapidly dissipate momentum within dense canopies [@raupachAveragingProceduresFlow1982; @baldocchiSpectralLagcorrelationAnalysis1988].
Within canopies, however, transport often fails to mimic diffusion – e.g., transport of momentum [@shawSecondaryWindSpeed1977] or heat [@raupachLagrangianAnalysisScalar1987] may occur in the direction opposite to that predicted by gradients of wind speed or temperature, respectively – making vertical wind profiles difficult to generalize from first principles [@denmeadScalarTransportPlant1987; @raupachApplyingLagrangianFluid1989; @meyersModellingPlantCanopy1987; @katulModelingCO2Sources1999; @harmanSimpleUnifiedTheory2007].
Such counter-gradient transport may arise from the intermittent generation, at the canopy surface, of large wakes or coherent eddy structures that periodically dip down through the canopy, gathering packets of warmed, humidified and CO~2~-enriched or depleted air from beneath and within the canopy and flushing them to the atmosphere above [@finniganTurbulenceWavingWheat1979; @baldocchiTraceGasExchange1991]. 
These "sweep-eject" events can couple understory conditions more directly to conditions above the canopy than beneath.
